A Generalized Visbility to Power-Spectrum Framework for 21\,cm Inteferometric Probes of Reionizations

### 2017-07-07 Rotation paper/draft_working.tex
save

%\maketitle
\author{
Yunfan Gerry Zhang \altaffilmark{1},
Adrian C. Liu \altaffilmark{1, 2},
Aaron R. Parsons\altaffilmark{1, 2}
}

@@

\begin{abstract}
The observational effort to measure the cosmological 21\,cm power spectrum with radio interferometers requires high sensitivity. Current visibility-based power spectrum pipelines, though shown to ease control of systematics, lack the ability to include partial redundancy. We introduce a method to include partial redundancy in such power spectrum pipelines of drift-scan arrays. Our method efficiently finds pairs of baselines to cross-multiply, and quantifies the sensitivity contributions of given baselines. 
Using the 128-element configurations and beams of the Precision Array for Probing the Epoch of Reionization (PAPER-128), as well as 4 planned versions of the Hydrogen Epoch of Reionization Array (HERA), we illustrate how our method applies to different arrays and predict the sensitivity improvements of including each baseline pairs. We show that inclusion of partially redundant baselines 
would account for $20\%$ to $60\%$ of the sensitivity of PAPER-128 and different configurations of HERA. 
\end{abstract}

@@

In the past, observational studies of reionization, including Gunn-Peterson measurements of quasi-stellar objects \citep{Fan2006} and cosmic microwave background (CMB) measurements of temperature and anisotropy \citep{Planck2016}, Kinetic Sunyaev-Zeldo'vich effect \citep{kszpatchy} and and Lyman alpha emitter clustering \citep{mcquinnLyA} have given us indications of the rough time-frame of reionization, but only limited constrains on the finer spatial and temporal variations. A surge of recent radio-astronomical experiments of reionization focus on measurement of the ``spin-flip'' transition of neutral
hydrogen of characteristic wavelength 21\,cm \citep{Furlanetto2006181,PritchardLoeb}.
In complement to the aforementioned probes, the 21\,cm brightness temperature distribution is a direct tracer of neutral hydrogen through the epoch of reionization, and thus tomography of the 21\,cm line is a direct measurement of the full temporal and spatial variations of this event. 
Before possible realization of full scale 21\,cm tomography, many current radio interferometric efforts
prioritize measuring the spatial power spectrum of 21\,cm brightness temperature fluctuations.
Current-generation instruments include the Precision Array for Probing
the Epoch of Reionization (PAPER) \citep{Ali2015,paper32}, Murchison
Widefield Array (MWA) \citep{Bowman2013, Tingay2013}, Low Frequency Array (LOFAR) \citep{LOFAR}. Next-generation instruments include the Hydrogen Epoch of Reionization
Array (HERA) (e.g. \citealt{HERA,HERAconfiguration,HERABEAM1,HERADISH2})  under construction, 
and the Square Kilometer Array Low (SKA-low) (e.g. \citealt{SKA1}), which is currently in planning.

The highly redshifted 21\,cm signal is faint and diffuse, in contrast to the localized bright sources that have been of interest in many traditional applications of radio astronomy.  Current experiments are sensitivity-starved, with the sensitivity requirement further increased considering the foreground contaminations 5 orders of magnitude brighter than the cosmological signal of interest. Thus modern low-frequency radio interferometers aimed at measuring the 21\,cm signal bears designs different from traditional instruments. Many specially designed arrays, such as PAPER and HERA, feature multiple copies of the same baselines to repeatedly measure the same Fourier signal to increase sensitivity. To satisfy the sensitivity need, modern arrays are large. Ranging upwards from 100 elements, these arrays are typically of the drift-scan type to limit cost.

Analysis pipelines for the 21\,cm power spectrum typically fall into two categories. In the first, more traditional technique, images are formed in Fourier domain through rotation synthesis, followed by a foreground mitigation step to construct a power spectrum. An alternative technique works directly with visibilities from baselines and cross multiplies them to form the power spectrum. Tracking closely the instrumental output of visibilities, this technique allows one to deal with systematics such as Radio-Frequency-Interference(RFI) and instrumental cross-talks in a more native context, and thus tend to provide good limits on the power spectrum. An example of the visibility-based pipeline was presented in \cite{Ali2015}, which provides the newest upper limit to the power spectrum
measurements with the 64-element version of PAPER (henceforth as PAPER-64). However, one disadvantage of existing visibility-based pipelines is the lack of use of partial redundancy. Baselines of different length and orientation still contain partially redundant information. While an imaging based power-spectrum pipeline naturally includes all redundancy information, visibility-based pipelines until now only cross multiply fully redundant baselines, in other words baselines of the same length and orientation.

The use of partially redundant baselines is in principle not a fundamental limitation of visibility-based pipelines. In fact, most instrumental reports to date do include partial redundancy. Recently, \cite{wterm} proposed a visibility-based approach to extract power spectrum from partially redundant baselines in tracking measurements. Arrays capable of tracking include MWA and LOFAR. Our work parallels this effort by focusing on drift-scan arrays that do not have tracking capabilities, such as PAPER, HERA and potentially SKA.
The Earth's rotation causes the baselines in a drift-scan array to pick up different modes of the sky with time. Rotation-synthesis makes use of the rotation-induced $uv$ coverage map to form image. In a visibility based pipeline, we can extract the same rotation-induced redundancy. In this point of view, baselines that are slightly different in length and orientation
``rotate into'' each other at a time delay. We can thus cross-multiply time-shifted visibilities, with the proper weighting, to form power spectra.  Due to the large number of elements of modern arrays (upward of 100), the task of cross-multiplying every baseline against every other, scaling as number of array elements to the fourth power, can be computationally formidable, and many pairs of baselines provide only negligible redundancy information. Our contribution is thus twofold. First we introduce a formalism to include pairs of partially-redundant baselines in a visibility based power spectrum pipeline. Secondly we show how to use the formalism to automatically pre-select baseline pairs and time offsets, making the problem computationally efficient. More precisely, our formalism allows one to simultaneously identify the baselines that give good
redundancy, find the time offset that corresponds to maximal redundancy for a given pair of baselines, and quantify the sensitivity associated with cross multiplying
such a pair of partially redundant baselines, which in turns is used as weight to combine measurements in a power spectrum pipeline.

@@

The rest of this paper is organized as follows. In section 2 we introduce some terminology and notations to be used for the rest of the paper. In section
3 we introduce the formalism for weighting partially redundant baselines.
In section 4 we present numerical tests of
this technique as well as the expected sensitivity improvement
with this method for HERA and PAPER-128 and with section 5 we conclude.

@@

In order to eliminate confusion and ambiguity for the rest of this paper, we
introduce some terminology that may differ from what is commonly found in the literature.

We make the distinction of a \textit{baseline}, which corresponds to two specific antennas, and a \textit{class of baselines}, which refers to all baselines of the same length and orientation in a given array. 
Baselines of the same class are traditionally called
``redundant baselines'', because they measure the same Fourier mode
in the sky.  We shall call baselines in the same class \textit{equivalent baselines}, and reserve the word \textit{redundancy} of two baselines in reference to a variable function of the relative time-offsets of their visibility time series. With this terminology, two equivalent
baselines are fully redundant with each other simultaneously at all
times. Non-equivalent baselines also have partial redundancy, and the redundancy can be maximized if their
respected time series are shifted with respect to one another by some delay. In other words, one baseline can be ``rotated into''
another if they are \textit{near-equivalent}.

\begin{figure}[H]
\includegraphics[width=\linewidth]{antpos128}

\caption{The PAPER128 layout. Each dot corresponds to the location of
an antenna. Top panel shows the antenna positions drawn to scale;
bottom panel show the antenna labels and distances of the 112-element grid, in other words excluding the outrigger antennas.
The numbering of the antennas in the bottom panel are original labels
during instrument assembly and does not bear significant
meaning. In the bottom panel, the two baselines (5,77) and (40,85) marked with red segments are example of an equivalent pair, both with separation denoted by \{1,0\}, for the
antennas are separated by 1 unit east and 0 unit north. Similarly,
the baselines marked in blue and yellow are examples
of the classes \{1,1\} and \{1,-1\}, respectively. Due to the small North-South separation within the grid, \{1,1\} is expected to be near-equivalent to \{1,0\}. }
\label{fig:AntPos}
\end{figure}

We shall use the 128-element PAPER array (henceforth as PAPER-128) to motivate our formalism and demonstrate our method, and extend our results to several HERA configurations in Sec \ref{sec:arrconf}. 
The PAPER array is located in the Karoo desert in South Africa (30:43:17.5
S, 21:25:41.8 E). The layout pattern with antenna labels are show
in Fig. \ref{fig:AntPos}. The array consists of a 112-element grid and 16 "outriggers" used primarily to aid calibration.  In the bottom panel, the two baselines (5, 77) and (40, 85), marked in red, are an example of equivalent pair. We denote a equivalency-class of baselines in the PAPER grid by their separations, in this case  \{1,0\}, for the
antennas are separated by 1 unit east and 0 unit north. Similarly,
the baselines marked in blue and yellow are respectively examples
of \{1,1\} and \{1, -1\}.
Note \{1,0\} and \{-1, 0\} for example are the same class and should
not be counted twice. Antennas in purely north-south baselines
are close (4m), and hence these baselines are not suitable
for cross-multiplication due to cross-talks. On the other hand, the close North-South separation means that classes such as \{1,0\} and  \{1,1\} are expected to be near-equivalent. The PAPER-64 analysis of \cite{Ali2015} used three classes of baselines, the PAPER-128
equivalent of which are 
\{2, 0\}, \{2, 1\} and \{2, -1\}. There each of these classes
of baselines are cross multiplied within itself. This paper provides the method for inter-class multiplications. To do so we will use the short hand notation $\{m,n\}:\{m',n'\}$ to denote a pair of baseline classes to be cross-multiplied.

@@

@@

Given a point source on the sky, each baseline maps to a point in $uvw$ space. As the Earth rotates with respect to
the source, the points trace out tracks in the $uvw$ space. 
We show in Fig. \ref{fig:Tracks} $uvw$ tracks of PAPER-128 over 12 hours, at 0.15GHz, for a source that passes through zenith.

\begin{figure}[H]
\includegraphics[width=\linewidth]{tracks3D}
\caption{Tracks of PAPER-128 (only a randomly selected subset of 32 baselines are shown) for frequency $\nu=0.15\text{GHz}$ and a hypothetical source that passes through zenith.
These tracks are traced out over 0.5 sidereal days. Color represents different baselines. 
As the Earth rotates, tracks are traced out counterclockwise. Tracks in $uvw$ space do not cross, even though the $uv$ projections do. }
\label{fig:Tracks}
\end{figure}

Traditionally, we can identify
redundancy of near-equivalent baselines as crossings
of the $uv$ tracks, a 2 dimensional projection of Fig. \ref{fig:Tracks}. As we see in Fig. \ref{fig:Tracks}, there would be many such crossings. However, there are several reasons that $uv$ track-crossing do not imply perfect redundancy. In fact, in our case $uv$ track-crossings are
not accurate enough for time offset determination, nor can it give estimate of the degree of redundancy.  The most obvious reason is that the 3 dimensional $uvw$ tracks do not actually cross. For drift-scan arrays, even a crossing in $uvw$ space does not imply perfect redundancy. This shall become evident in the next section, after we develop a more general formalism that accounts for the point spread function of the finite beams, and estimate the degree of redundancy for general combination of baselines at a general time-offset. An explicit explanation of track-crossing will be presented in Appendix \ref{sec:appA}.

\subsection{Formalism}
In this section we formulate the relation between 
the product of visibilities from two arbitrary baselines and the power spectrum.

We begin with the visibility as commonly defined in the literature (e.g.
\citealt{first-paper}): 
\begin{equation}
\begin{aligned}V_{\nu}(\boldsymbol{b}) & =\int d\Omega A(\hat{\boldsymbol{s}},\nu)\phi(\nu)I_{\nu}(\hat{\boldsymbol{s}})\exp\left(-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right),\\
 & \approx\frac{2k_{B}}{\lambda^{2}}\int d\Omega A(\hat{\boldsymbol{s}},\nu)\phi(\nu)T(\hat{\boldsymbol{s}})\exp\left(-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right),
\end{aligned}
\label{eq:Vis1}
\end{equation}
Here $k_B$ is Boltzmann's constant, $\lambda$ is the mean wavelength, $\boldsymbol{b}$ is the baseline length, $\hat{\boldsymbol{s}}$ and $\Omega$ are a direction in the
sky and its corresponding solid angle. Inside the integral we have $\phi(\nu)$ the frequency bandpass profile, $A_{\nu}$ the (frequency
dependent) primary beam, and $I$ the specific intensity, which has been
related to $T$, the brightness temperature in the Rayleigh-Jeans
limit. The beam reception pattern $A$ is dimensionless,
normalized to 1 at its peak (zenith), and we assume it to be the same
for all baselines. We show sample beams of HERA and PAPER antennas in Fig. \ref{fig:Beam} for reference.

@@

In practice,
power spectrum measurements are typically taken from a few ten MHz centered around the corresponding redshift of interest (e.g. 150 MHz for z=9.5).

Delayed-transformed visibility has been widely adopted in recent years since it has been shown that foregrounds are isolated in delay space. We define the delay-transformed visibility by Fourier transforming visibility along the frequency axis\citep{delay-transform} as:
\small
\begin{equation}
\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left(-2\pi i\nu\tau\right),\\
 & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{\boldsymbol{s}},\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}+\tau\right)\right]. 
\end{aligned}
\label{eq:Vb1}
\end{equation}
\normalsize
Here the delay $\tau$ is the Fourier variable. Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
an integral over observational coordinates $\hat{\boldsymbol{s}}$ and $\nu$. For notational simplicity here and in the rest of this paper, we shall absorb the bandpass $\phi$ inside the primary beam $A$.

@@

In the thin-shell limit, we can write:
\begin{equation}
d^2r=X^2d\Omega. 
\label{eq:thinshell}
\end{equation}
Note that in Eq.\ref{eq:thinshell} no flat-sky approximation has been made; the angular integral is still performed over the dome.

@@

\begin{equation}
\begin{aligned} & \langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau)\rangle\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)|\\
 & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)\right].\end{aligned}
\label{eq:final}
\end{equation}

In other words the power spectrum estimate from visibilities of a baseline pair is given by 
\begin{equation}
 P(k_{b,\tau}) = \left(\frac{\lambda^{2}}{2k_{B}}\right)^{2} \frac{\langle V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau)\rangle}{\Theta}, 
 \label{eq:opp}
\end{equation}
where the weight
\begin{equation}
\Theta \equiv\int d\nu \Theta_{\nu}, 
\label{eq:Theta}
\end{equation}
and

\[
\Theta_{\nu} \equiv \int \frac{d\Omega}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)| e^{-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}e^{i\psi_{\nu}}.
\label{eq:Thetanu}
\]

Notice that $\Theta$ has no dependence on $\tau$. In Eq. \ref{eq:Thetanu} we introduced an extra phase $\psi$, the origin of which we shall explain in \ref{sec:rephs}. We point out that although all our derivations focused on drift-scan telescopes, we can get the analogous result for tracking measurements simply by noticing that for a tracking primary beam, $\Gamma$ becomes a rotation around zenith, and so:
\[
\Theta_{\nu} \equiv \int \frac{d\Omega}{X^{2}Y}|A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\boldsymbol{s}},\nu)| e^{-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}e^{i\psi_{\nu}}.
\label{eq:Thetanu_tracking}
\]

>>>>

%\maketitle
\author{
Yunfan Gerry Zhang \altaffilmark{1},
Adrian Liu \altaffilmark{1, 2},
Aaron R. Parsons\altaffilmark{1, 2}
}

@@

\begin{abstract}
Measuring the cosmological 21\,cm power spectrum with radio interferometers requires high sensitivity. Many modern low-frequency interferometers are thus designed with antennas placed on regular grid to maximize redundancy. Current visibility-based power spectrum pipelines, though shown to ease control of systematics, lack the ability to include partial redundancy. We introduce a method to include partial redundancy in such power spectrum pipelines of drift-scan arrays. Our method pairs baselines to cross-multiply at a time lag, and quantifies the sensitivity contributions of each pair of baselines. 
Using the 128-element configurations and beams of the Donald C. Backer Precision Array for Probing the Epoch of Reionization (PAPER-128), and 4 planned versions of the Hydrogen Epoch of Reionization Array (HERA), we illustrate how our method applies to different arrays and predict the sensitivity improvements of including each baseline pairs. We show that inclusion of partially redundant baselines 
would account for $20\%$ to $60\%$ of the sensitivity of PAPER-128 and different configurations of HERA. 
\end{abstract}

@@

Observational studies of reionization, including Gunn-Peterson measurements of quasi-stellar objects \citep{Fan2006} and cosmic microwave background temperature and anisotropy measurements (CMB;\citealt{Planck2016}), the Kinetic Sunyaev-Zeldo'vich effect \citep{kszpatchy} and Lyman alpha emitter clustering \citep{mcquinnLyA} have given us indications of the rough time-frame of reionization, but only limited constraints on the finer spatial and temporal structures. A surge of recent radio-astronomical experiments of reionization focus on measuring the ``spin-flip'' transition of neutral
hydrogen of characteristic wavelength 21\,cm \citep{Furlanetto2006181,PritchardLoeb}.
The 21\,cm brightness temperature is a direct tracer of neutral hydrogen through the epoch of reionization, and thus tomography of the 21\,cm line is a direct measurement of the full temporal and spatial variations of this event. 
Before realizing of full-scale 21\,cm tomography, many current radio interferometric efforts
aim to measure the spatial power spectrum of 21\,cm brightness temperature fluctuations.
Current-generation instruments include the Donald C. Backer Precision Array for Probing
the Epoch of Reionization (PAPER; \citealt{Ali2015,paper32}), Murchison
Widefield Array (MWA; \citealt{Bowman2013, Tingay2013}), Low Frequency Array (LOFAR; \citealt{LOFAR}). Next-generation instruments include the Hydrogen Epoch of Reionization
Array (HERA; e.g. \citealt{HERA,HERAconfiguration,HERABEAM1,HERADISH2})  under construction, 
and the Square Kilometer Array Low (SKA-low; e.g. \citealt{SKA1}) currently in planning stages.

The highly redshifted 21\,cm signal is faint and diffuse, in contrast to the localized bright sources targeted by many traditional radio telescopes.  Current 21\,cm experiments are sensitivity-starved, with the sensitivity requirement further increased considering the foreground contaminations 5 orders of magnitude brighter than the cosmological signal of interest. Low-frequency radio interferometers aiming to measure the 21\,cm signal are thus designed differently from traditional instruments. Specially designed arrays such as PAPER and HERA feature multiple copies of the same baselines to repeatedly measure the same Fourier signal to increase sensitivity \citep{first-paper}. To satisfy the sensitivity need, modern arrays are large. Ranging upwards from 100 elements, these arrays are typically of the drift-scan type to limit cost.

Analysis pipelines for the 21\,cm power spectrum typically fall into two categories. In the first, images are formed in Fourier domain through rotation synthesis, followed by a foreground mitigation step to construct a power spectrum. An alternative technique works directly with visibilities from baselines, delay-transforming and cross-multiplying them to form the power spectrum. This technique avoids many systematics associated with combining data from different baselines and tracks the native sampling of the interferometer. An example of the visibility-based pipeline was presented in \cite{Ali2015}, which provides the newest upper limit to the power spectrum
measurements with the 64-element version of PAPER (henceforth as PAPER-64). However, one disadvantage of existing visibility-based pipelines is their lack of use of partial redundancy. Baselines of different lengths and orientations contain partially redundant information. While imaging based power-spectrum pipelines naturally includes all redundancy information, visibility-based pipelines so far only cross multiply fully redundant baselines, i.e. baselines of the same length and orientation.

Visibility-based pipelines are not fundamentally limited from using partially redundant baselines. In fact, most sensitivity forecasts to date do include partial redundancy. Recently, \cite{wterm} proposed a visibility-based approach to extract power spectrum from partially redundant baselines in sky-tracking measurements. Arrays capable of such tracking include MWA and LOFAR. Our work parallels this effort by focusing on drift-scan arrays that do not have tracking capabilities, such as PAPER, HERA and potentially SKA-low.

The Earth's rotation causes the baselines in a drift-scan array to pick up different modes of the sky with time. Rotation-synthesis makes use of the rotation-induced $uv$ coverage map to form image. In a visibility based pipeline, we can extract the same rotation-induced redundancy. In this point of view, baselines that are slightly different in length and orientation
``rotate into'' each other at a time delay. We can thus cross-multiply time-shifted visibilities, with the proper weighting, to form power spectra.  Due to the large number of elements of modern arrays (upward of 100), the task of cross-multiplying every baseline against every other, scaling as number of array elements to the fourth power, can be computationally formidable, and many pairs of baselines provide only negligible redundancy information. Our contribution is thus twofold. First we introduce a formalism to include pairs of partially-redundant baselines in a visibility based power spectrum pipeline. Secondly we show how to use the formalism to automatically pre-select baseline pairs and time offsets, making the problem computationally efficient. More precisely, our formalism allows one to simultaneously identify the baselines that give good
redundancy, find the time offset that corresponds to maximal redundancy for a given pair of baselines, and quantify the sensitivity associated with cross multiplying
such a pair of partially redundant baselines, which in turns is used as weight to combine measurements in a power spectrum pipeline.

@@

The rest of this paper is organized as follows. In section 2 we introduce some terminology and notation used in the rest of the paper. In section
3 we introduce the formalism for weighting partially redundant baselines.
In section 4 we present numerical tests of
this technique as well as the expected sensitivity improvement
 this method provides for HERA and PAPER-128 pipelines, and with section 5 we conclude.

@@

In order to avoid confusion and ambiguity for the rest of this paper, we
introduce some terminology that may differ from what is commonly found in the literature.

We make the distinction between a \textit{baseline}, which corresponds to two specific antennas, and a \textit{class of baselines}, which refers to all baselines of the same length and orientation in a given array. 
Baselines of the same class are traditionally called
``redundant baselines'', because they measure the same Fourier mode
in the sky.  We shall call baselines in the same class \textit{equivalent baselines}, and reserve the word \textit{redundancy} of two baselines in reference to a variable function of the relative time-offsets of their visibility time series. With this terminology, two equivalent
baselines are fully redundant with each other simultaneously at all
times. Non-equivalent baselines also have partial redundancy, and the redundancy can be maximized if their
respected time series are shifted with respect to one another by some delay.

\begin{figure}[H]
\includegraphics[width=\linewidth]{antconfig}

\caption{The PAPER128 layout. Each dot corresponds to the location of
an antenna. The top panel shows the antenna positions drawn to scale;
the bottom panel show the antenna labels and distances of the 112-element grid, in other words excluding the outrigger antennas.
The numbering of the antennas in the bottom panel are original labels
during instrument assembly and does not bear significant
meaning. In the bottom panel, the two baselines marked with red segments are example of an equivalent pair, each with antennas separated by 1 unit east and 0 unit north. Similarly,
the baselines marked in blue and yellow are examples
of classes that , due to the small North-South separation within the grid (see top panel), are expected to be nearly equivalent to the red baselines. }
\label{fig:AntPos}
\end{figure}

We shall use the 128-element PAPER array (henceforth as PAPER-128) to motivate our formalism and demonstrate our method, and extend our results to several HERA configurations in Sec \ref{sec:arrconf}. 
The PAPER array is located in the Karoo desert in South Africa (30:43:17.5
S, 21:25:41.8 E). The layout pattern with antenna labels is shown
in Fig. \ref{fig:AntPos}. The array consists of a 112-element core is a rectangular grid, and 16 "outriggers" used primarily to aid calibration.  In the bottom panel, the two baselines marked in red, are an example of equivalent pair. We denote a equivalency-class of baselines in the PAPER grid by their separations, in this case  \{1,0\}, for the
antennas are separated by 1 unit east and 0 unit north. Similarly,
the baselines marked in blue and yellow are respectively examples
of \{1,1\} and \{1, -1\}.
Note \{1,0\} and \{-1, 0\} for example are the same class and should
not be counted twice. Antennas in purely north-south baselines
are close (4m), and hence these baselines are not suitable
for cross-multiplication due to cross-coupling. On the other hand, the close North-South separation means that classes such as \{1,0\} and  \{1,1\} are expected to be near-equivalent. The PAPER-64 analysis of \cite{Ali2015} used three classes of baselines, the PAPER-128
equivalent of which are 
\{2, 0\}, \{2, 1\} and \{2, -1\}. There each of these classes
of baselines are cross multiplied within itself. This paper provides the method for inter-class multiplications. To do so we will use the short hand notation $\{m,n\}:\{m',n'\}$ to denote a pair of baseline classes to be cross-multiplied.

@@

In this section we introduce our method to cross-multiply nearly equivalent baseline classes.

@@

Relative to a phase center on the sky, each baseline maps to a point in $uvw$ space. As the Earth rotates with respect to
the source, the points trace out tracks in the $uvw$ space. 
We show in Fig. \ref{fig:Tracks} $uvw$ tracks of PAPER-128 over 12 sidereal hours, at 0.15GHz, relative to a phase center that passes through zenith.

\begin{figure}[H]
\includegraphics[width=\linewidth]{rotation}
\caption{Tracks of PAPER-128 for a selection of 16 baselines and frequency $\nu=0.15\text{GHz}$ and a phase center that passes through zenith.
Tracks are traced out over half a sidereal day. Color represents different baselines. 
As the Earth rotates, tracks are traced out counterclockwise. Tracks in $uvw$ space do not cross, even though the $uv$ projections do. }
\label{fig:Tracks}
\end{figure}

Equivalent baselines follow identical $uvw$ tracks. Traditionally, we can identify
redundancy of nearly equivalent baselines as crossings
of the $uv$ tracks, a 2 dimensional projection of Fig. \ref{fig:Tracks}. As we see in Fig. \ref{fig:Tracks}, there would be many such crossings. However, there are several reasons that $uv$ track-crossing do not imply perfect redundancy. In fact, in our case $uv$ track-crossings are
not accurate enough for time offset determination, nor can it give estimate of the degree of redundancy.  The most obvious reason is that the 3 dimensional $uvw$ tracks do not actually cross. For drift-scan arrays, even a hypothetical crossing in $uvw$ space would not imply perfect redundancy. This shall become evident in the next section, after we develop a more general formalism that accounts for the point spread function of the finite beams, and estimate the degree of redundancy for general combination of baselines at a general time-offset. Then, relation of track-crossing and redundancy for both drift-scan and tracking measurements will be explored in Appendix \ref{sec:appA}.

\subsection{Formalism}
In this section we formulate the relation between 
the product of delay-transformed visibilities from two arbitrary baselines and the power spectrum.

We begin with the visibility as commonly defined in the literature (e.g.
\citealt{first-paper}): 
\begin{equation}
\begin{aligned}V_{\nu}(\boldsymbol{b}) & =\int d\Omega A(\hat{\boldsymbol{s}},\nu)\phi(\nu)I_{\nu}(\hat{\boldsymbol{s}})\exp\left(-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right),\\
 & \approx\frac{2k_{B}}{\lambda^{2}}\int d\Omega A(\hat{\boldsymbol{s}},\nu)\phi(\nu)T(\hat{\boldsymbol{s}})\exp\left(-2\pi i\frac{\nu}{c}\boldsymbol{b}\cdot\hat{\boldsymbol{s}}\right),
\end{aligned}
\label{eq:Vis1}
\end{equation}
where $k_B$ is Boltzmann's constant, $\lambda$ is the mean wavelength, $\boldsymbol{b}$ is the baseline vector, $\hat{\boldsymbol{s}}$ and $\Omega$ are a direction in the
sky and its corresponding solid angle. Inside the integral we have $\phi(\nu)$ the frequency bandpass profile, $A_{\nu}$ the (frequency-dependent) primary beam, and $I$ the specific intensity, which has been
related to $T$, the brightness temperature in the Rayleigh-Jeans
limit. The beam power pattern $A$ is dimensionless,
normalized to 1 at its peak (zenith), and we assume it to be the same
for all baselines. We show sample beams of HERA and PAPER antennas in Fig. \ref{fig:Beam} for reference.

@@

Power-spectrum measurements are typically taken from $\sim10$MHz centered around the corresponding redshift of interest (e.g. 150 MHz for z=9.5).

Delayed-transformed visibility has gained popularity in recent years since it has been shown that foregrounds are isolated in delay space. We define the delay-transformed visibility by Fourier transforming visibility along the frequency axis \citep{delay-transform} as:
\small
\begin{equation}
\begin{aligned}V(\boldsymbol{b},\tau) & =\int d\nu V_{\nu}(\boldsymbol{b})\phi(\nu)\exp\left(-2\pi i\nu\tau\right),\\
 & =\frac{2k_{B}}{\lambda^{2}}\int d\Omega d\nu A(\hat{\boldsymbol{s}},\nu)T(\hat{\boldsymbol{s}},\nu)\exp\left[-2\pi i\nu\left(\frac{\boldsymbol{b}\cdot\hat{\boldsymbol{s}}}{c}+\tau\right)\right]. 
\end{aligned}
\label{eq:Vb1}
\end{equation}
\normalsize
Here the delay $\tau$ is the Fourier dual of $\nu$. Eq. \eqref{eq:Vb1} expresses the delay-transformed visibility as
an integral over observational coordinates $\hat{\boldsymbol{s}}$ and $\nu$. For notational simplicity here and in the rest of this paper, we shall absorb the bandpass $\phi$ inside the primary beam $A$.

@@

In the thin-shell limit, we can write:
\begin{equation}
d^2r=X^2d\Omega. 
\label{eq:thinshell}
\end{equation}
Note Eq. \eqref{eq:thinshell} does not require the flat-sky approximation; the angular integral is still performed over the curved sky.

@@

\begin{equation}
\begin{aligned} & \langle e^{i\psi_{\nu}} V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau)\rangle\\
 & =\left(\frac{2k_{B}}{\lambda^{2}}\right)^{2}P(k_{b,\tau})\int\frac{d\Omega d\nu}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu)\\
 & \qquad \qquad \qquad \qquad \exp\left[-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)+i\psi_{\nu}\right].
 \end{aligned}
\label{eq:final}
\end{equation}

In other words the power spectrum estimate from visibilities of a baseline pair is given by 
\begin{equation}
 P(k_{b,\tau}) = \left(\frac{\lambda^{2}}{2k_{B}}\right)^{2} \frac{\langle e^{i\psi_{\nu}} V^{*}(\boldsymbol{b},\tau)V(\boldsymbol{b'},\tau)\rangle}{\Theta}, 
 \label{eq:opp}
\end{equation}
where the weight is defined as 
\begin{equation}
\Theta \equiv\int d\nu \Theta_{\nu}, 
\label{eq:Theta}
\end{equation}
with

\[
\Theta_{\nu} \equiv \int \frac{d\Omega}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\Gamma\hat{\boldsymbol{s}},\nu) e^{-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}e^{i\psi_{\nu}}.
\label{eq:Thetanu}
\]

Notice that $\Theta$ has no dependence on $\tau$. In Eq. \eqref{eq:Thetanu} we introduced an extra phase $\psi$, the origin of which we shall explain in \ref{sec:rephs}. We point out that although all our derivations focused on drift-scan telescopes, we can get the analogous result for tracking measurements simply by noticing that for a tracking primary beam, $\Gamma$ becomes a rotation around zenith, and so:
\[
\Theta_{\nu} \equiv \int \frac{d\Omega}{X^{2}Y}A^{*}(\hat{\boldsymbol{s}},\nu)A(\hat{\boldsymbol{s}},\nu) e^{-i2\pi\frac{\nu}{c}\left(\hat{\boldsymbol{s}}\cdot\boldsymbol{b}-\Gamma\hat{\boldsymbol{s}}\cdot\boldsymbol{b'}\right)}e^{i\psi_{\nu}}.
\label{eq:Thetanu_tracking}
\]

Roughly speaking, Eqs. \eqref{eq:opp} through \eqref{eq:Thetanu} tell us that the cross multiplications of visibilities at a time delay is proportional to the power spectrum times the Fourier
transform of the cross multiplied beam patterns. As a check, when applied to equivalent baselines,
$\boldsymbol{b}=\boldsymbol{b'}$, $\hat{\boldsymbol{s}}=\Gamma\hat{\boldsymbol{s}}$, and Eq.(\ref{eq:final}) reduces to Eq.(B9) of \cite{paper32}. 
With Eq. \eqref{eq:opp} and Eq. \eqref{eq:Theta} we can, for any given pair of baseline classes and time delay, estimate the degree of redundancy, here represented by $\Theta$. This allows us to achieve our goals stated in the introduction: to identify 
candidate baseline pairs with significant redundancy, to find the time offset that maximizes redundancy, and to quantify the degree of such redundancy. We can do all the above simply by computing the weight $\Theta$ from
Eq.(\ref{eq:Theta}) for various time offsets.

\subsection{Rephasing \label{sec:rephs}}
If $\psi_{\nu}=0$ in  Eq. \eqref{eq:Thetanu}, $\Theta_{\nu}$ at the peak of correlation is in general complex, and often far from real. Furthermore, this phase of peak correlation is inevitably frequency dependent. This frequency dependence of the phase would lead to destructive interference when we integrate over frequency, unless we correct $\Theta_{\nu}$ by an extra phase $\psi_\nu$.
